---
title: 'STAA57 Final Project Report '
author: "Brian Lin, Ian Gu, Minyi Ma, Navin Chandradat"
output:
  html_document:
    df_print: paged
---
```{r, echo=FALSE, message=FALSE, warning=FALSE}
## echo=FALSE stops the code chunk from appearing in the html document
## warning=FALSE / message=FALSE stop R warnings/messages from appearing in the knit document
library(tidyverse)
```


***

<!-- 
(everything below in here is an .html comment)
Instructions: 
- Use typical structure of data analysis report, w/ sections:
    + Introduction
    + Data
    + Analysis
    + Summary
    + Appendix (optional)
- Report should be 5 pages MAX (when compiled/knit as pdf)
- Appendix does not count towards page limit
- Put all your data in a separate "./data/" sub-folder 
- Submit compressed project folder 
- You submissions should be reproducible (i.e. I should be able to   
  re-compile your report with a click of the Preview/Knit button) 
-->

## Introduction 

(*Description of questions that are being investigated*)

### How fast will you be able to recoup your investment in post-secondary education - money approach?

>Our first research question addresses the question of how fast one is able to recoup their investment in post-secondary education. To investigate this question, analysis will be done on the income of those who graduate out of high school and enter the labour force, as well as the income of those who invest in a post-secondary education and begin working. 

### Education Level vs. Quality of Life - future progress approach:
>Our second research question addresses the quality of life of different people, in comparison to their education level. This question hopes to answer the "worth" part of the question "Is a university education *worth* the investment?", as it compares the quality of life between people of different education levels. One of the main concerns of this question is the question of how to measure quality of life within the variables readily available, or in other words, how to quantify it in a way that can be used to analyze statistically without access to data such as happiness, emotional well-being, and such. The first aspect is considering their ability to tolerant stress，for this part we analyze the unemployment rate and length of unemployment period of people who attend univeristy or not. In the aim to see whether post-secondary education lead them a better future, which means will they tend to be more actively seeking and more eaiser to get job in the job market when they face the condition of eagerly seeking for job. The second aspect is onsidering the life, we firstly analysis the marry condition of people at same age, which enable us to see whether post-secondary

## Data 

(*Description of data used to answer the questions*)

### Main data used:
* Lfs2013: Labour force survey, 2013
  + The Labour Force Survey is a monthly survey in which citizens of Canada are asked to respond to several questions which measure the state of the labour market in Canada. Among the information extracted from the survey are statistics like the employment status of citizens, the income of their jobs etc.

* Cis2013: Canadian Income Survey, 2013
  + The Canadian Income Survey provides information related to each surveyed Canadian citizen's job status such as their age group, education level, and unemployment time length. We will be using this data to extrapolate the unemployment condition.

##### Group by:
* educ90: Highest educational attainment (1990- )
  + This will be the grouping variable that helps identify whether people finish university in lfs.
* age_12: Five-year age group of respondent in lfs
  + This will be the grouping variable that shows unemployment rate and income difference through different age groups.
* HLEV2G:  Highest level of education
  + This will be the grouping variable that helps identify whether people finish university in cis.


##### Exclusion:

* FLLPRTP: Full-time or part-time student during refyear
  + This binary variable will be used to filter out all current full-time students, to make sure they won't effect the unemployment rate.


##### Independent variables to consider:
For income:
* hrlyearn: Usual hourly wages
* age_12: Five-year age group of respondent in lfs
（for this variable mainly use the combination of hrlyearn and it）

For unemployment:
*lfsstat: Labour force status
* age_12: Five-year age group of respondent in lfs
（for this variable mainly use the combination of lfsstat and it
* WKSUEM: All persons aged 16 to 69 and persons aged 70 and over who worked during the reference year
 
## Analysis 
(*Main body of analysis, containing data summaries (tables/plots),  statistical analyses (estimations, tests), and modelling (regression, classification) *)
### Analysis upon income.
<!-- You can organize content in subsections -->
(*Plots should be created in R, and not imported from file*)

```{r}
ngs = read.csv("ngs.csv")
cis = read.csv("cis.csv")
lfs = read.csv("lfs.csv")
```


> In this part, we want to see if attending post secondary education, or specifically, university education can lead to a higher income.

```{r}
library(ggplot2)
#the highest education effect on 1990 later
lfs_90 = lfs %>% mutate(educ90 = factor(educ90, levels=0:6, labels = c("Not attend", "Not attend", "Not attend", "Not finshied", "Finished", "Finished", "Finished"))) %>% mutate(pick = is.na(educ90)) %>% filter(pick == FALSE) %>% select(1:6)
lfs_90 %>% filter(!is.na(hrlyearn)) %>% ggplot() + geom_boxplot(aes(x = educ90, y=hrlyearn)) + xlab("educ_status")  
```
> Firstly we do a basic comparison of the hourly earning between the highest education level. The data gives 7 different types of education level, from entry level to the Master degree. We choose the education data of 1990 later to make sure its up to date and devide these people into three groups showing the level they engage in the post secondary education. 
Based on the boxplot, it is clear that in average, most people who obatained a university degree are tend to have a higher hourly income. Also, the highest income that group of people can reach are much higher that two other groups. 
Another thing that can be discovered from the graph is that though highly closed, the *NF* group(people who does attend but not finish the post secondary education), however, have the lowest income on average, and the also the highest income of that group is the lowest among that of the three groups. Which may shows that the paid back of taking university eucation is also related to whether taking the complete 4 years or more. 
To make it easier to test for different age, we now combine NF and N(since their result are really close considering hrlyearn)

```{r}
lfs_90_N = lfs %>% mutate(educ90 = factor(educ90, levels=0:6, labels = c("N", "N", "N", "N", "T", "T", "T"))) %>% mutate(pick = is.na(educ90)) %>% filter(pick == FALSE) %>% select(1:6) %>% filter(educ90 == "N") 
lfs_90_T = lfs %>% mutate(educ90 = factor(educ90, levels=0:6, labels = c("N", "N", "N", "N", "T", "T", "T"))) %>% mutate(pick = is.na(educ90)) %>% filter(pick == FALSE) %>% select(1:6) %>% filter(educ90 == "T")
N_age = lfs_90_N %>% group_by(age_12) %>% summarise(avg_income_N = mean(hrlyearn, na.rm = TRUE))
T_age = lfs_90_T %>% group_by(age_12) %>% summarise(avg_income_T = mean(hrlyearn, na.rm = TRUE))
left_join(T_age, N_age, by = "age_12") %>% mutate(diff = avg_income_T - avg_income_N) %>% knitr::kable()
left_join(T_age, N_age, by = "age_12") %>% mutate(diff = avg_income_T - avg_income_N) %>% ggplot(aes(x = age_12, y = diff)) + geom_point( )
```

>Based on the data, we group the peoples age for 5 years each. It is clear that for all age group, owning a secondary eduaction degree is a guarantee of higher hourly earn.
For more specific apporach, we then look at the difference of the difference of hourly earning between people who have degree and who not. The difference keep increasing till it reach the highest pointat group10, age between 60-64. This is mainly showing that the effect of posy secondary degree will become larger as age grows. While the data decrease after group10, is mainly due to the retirement and also back to years ago the post-secondary are not that necessary in 20th century.



#### Analysis upon unemployment


> Secondly, we want to see if the university offer people higher chance to get a job, which means will the unemployment rate be lower or will the unemployment period will be shorter if for the population who does finsih their university education. 

```{r}
lfs_N = lfs %>% filter(lfsstat != 6) %>%mutate(educ90 = factor(educ90, levels=0:6, labels = c("N", "N", "N", "N", "T", "T", "T"))) %>% mutate(pick = is.na(educ90)) %>% filter(pick == FALSE) %>% select(1:6) %>% filter(educ90 == "N") %>% mutate(lfsstat = (lfsstat >= 3)) %>% group_by(age_12) %>% summarise(unemploy_rate_N = mean(lfsstat))
lfs_T = lfs %>% filter(lfsstat != 6) %>%mutate(educ90 = factor(educ90, levels=0:6, labels = c("N", "N", "N", "N", "T", "T", "T"))) %>% mutate(pick = is.na(educ90)) %>% filter(pick == FALSE) %>% select(1:6) %>% filter(educ90 == "T") %>% mutate(lfsstat = (lfsstat >= 3)) %>% group_by(age_12) %>% summarise(unemploy_rate_T = mean(lfsstat))
left_join(lfs_T,lfs_N, by = "age_12") %>% mutate(diff = unemploy_rate_N - unemploy_rate_T )%>% knitr::kable()
```

>We first estimate the unemployment rate for both group through age group. The "unemploy_rate_T" stands for the rate for people attend univeristy and "unemploy_rate_N" stands for those who not. Clearly, most of the diff is positive, which means that people who didn't have university degree will tend to have higher unemployment rate. Also, there is a trend showing that with ages grow, the difference tend to be smaller, which reveals that the university education are becoming more important as time goes. 

```{r}
cis_new = cis %>% filter(hlev2g %in% c(1,2,3,4)) %>% mutate(hlev2g = factor(hlev2g, levels=1:4, labels = c("Not Attend", "Not Attend", "Not Finished", "Attended"))) %>% filter(wksuem %in% c(0:52)) 
cis_new %>% filter(wksuem != 0) %>% ggplot(aes(x = wksuem, fill = hlev2g))+geom_histogram()
```

>Also, we use the data from cis to see if the post secondary degree would effect people on their unemployment length. Based on the graph above, though for almost each length there are both people with no degree and people with degree, but the people who don't have degree tend to have longer weeks of unemployment. 

```{r}
library(coin)

```

```{r}
library(ggmosaic)
lfs_90 %>% filter(age_12 %in% c(3:7))%>% mutate(marstat = factor(marstat, level = 1:6, labels = c("Married", "Living in common law", "Widowed", "Seprated", "Divorced", "Single, never married"))) %>% ggplot() + geom_mosaic(aes(x = product(educ90), fill = marstat)) + xlab("educ") + ylab("marstat")
```
>Finally, we want to see if having education have any effects on people's daily life. With the mosaic graph, we can easily see that the education can lead to a higher rate of married.

```{r}
lfs_90 %>% filter(age_12 %in% c(3:7)) %>% mutate(marstat = factor(marstat, level = 1:6, labels = c("Married", "Living in common law", "Widowed", "Seprated", "Divorced", "Single, never married"))) %>% independence_test(marstat ~ educ90, data = ., 
      alternative = "two.sided" )
```


## Summary 

>From our analysis we can see that in terms of hourly income, individuals that do opt for a university education end up earning a higher hourly wage. And if you were to project this by looking at the affect of this over time we see that the difference of wages of those who attended univeristy and those who did not attend grows and plateaus at around the age of 60-64, the typical age of retirement. Now in terms of actually securing a job after attending univeristy, we turn our analysis to the unemployment rates of those who attended univerisity and those who did not. Once we've done our calculations we see that there is indeed a difference in the respective unemployment rates. In fact they are indeed lower for the individuals who completed a university degree than those who chose not to. Now as time progresses we see that the difference does seem to get smaller and smaller but doesn't level out until after approximately the age of retirement. We also looked at the time of unemployment of individuals (in weeks) of the two respective groups. Our research showed us that individuals who completed university were umemployed for fewer weeks than those who did not complete a degree. Now we turn back to the initial question about "Is a university education worth it?". And based on the numbers for hourly wage and unemployment we can conclude that a university education is indeed worth the time, money and effort asked from young adults attending.   

```{r}
cis_new %>% filter(efmjsi %in% c(02, 03, 05, 06)) %>% filter(efmjie == 1) %>% filter(agegp %in% c(5:15))%>% mutate(agegp = factor(agegp, levels = 5:15, labels = c("18~19","20~24", "25~29", "30~34", "35~39", "40~44", "45~49", "50~54", "55~59", "60~64", "65~69")))%>% ggplot(aes(x = hlev2g, y= efmjie)) + geom_histogram(stat = "identity") + facet_wrap(.~agegp) + theme(axis.text.x = element_text(angle = 45, hjust = 1))
```




***  

## Appendix 

(*Everything here does not count towards page limit*)

