---
title: "STAA57 Final Project Report "
author: "Brian Lin, Ian Gu, Minyi Ma, Navin Chandradat"
output: html_notebook
---
```{r, echo=FALSE, message=FALSE, warning=FALSE}
## echo=FALSE stops the code chunk from appearing in the html document
## warning=FALSE / message=FALSE stop R warnings/messages from appearing in the knit document
library(tidyverse)
```


***

<!-- 
(everything below in here is an .html comment)

Instructions: 
- Use typical structure of data analysis report, w/ sections:
    + Introduction
    + Data
    + Analysis
    + Summary
    + Appendix (optional)
- Report should be 5 pages MAX (when compiled/knit as pdf)
- Appendix does not count towards page limit
- Put all your data in a separate "./data/" sub-folder 
- Submit compressed project folder 
- You submissions should be reproducible (i.e. I should be able to   
  re-compile your report with a click of the Preview/Knit button) 

-->

## Introduction 

(*Description of questions that are being investigated*)

###How fast will you be able to recoup your investment in post-secondary education?
>Our first research question addresses the question of how fast one is able to recoup their investment in post-secondary education. To investigate this question, analysis will be done on the income of those who graduate out of high school and enter the labour force, as well as the income of those who invest in a post-secondary education and begin working. Unemployment rate will also be considered when analyzing the results of the LFS.

###Education Level vs. Quality of Life:
>Our second research question addresses the quality of life of different people, in comparison to their education level. This question hopes to answer the "worth" part of the question "Is a university education *worth* the investment?", as it compares the quality of life between people of different education levels. One of the main concerns of this question is the question of how to measure quality of life within the variables readily available, or in other words, how to quantify it in a way that can be used to analyze statistically without access to data such as happiness, emotional well-being, and such.

## Data 

(*Description of data used to answer the questions*)

### How fast will you be able to recoup your investment in post-secondary education?

##### Key Variables:
* LFSTAT, labour force status in 2013 reference week. Used to filter and determine unemployment rate
* CNDPERSP, total personal income from all sources in 2013. Will be used with other variables to determine income.
* JOBINCP, gross annual earnings from 2013. Will be used with other variables to determine income.
* DBTALGRD, debt of all loans at time of graduation (2009/10). Will be used to determine how much investment someone has made in post-secondary education prior to beginning employment.
 
### Attempting to quantify Quality of Life:

##### Group by:
* age_12: Five-year age group of respondent
* HLEV2G:  Highest level of education

* HLEV2G:  Highest level of education
  + This will be the grouping variable that shows sorts the QOL of different deducation levels.

##### Exclusion:

* FLLPRTP: Flag - student during refyear
  + This binary variable will be used to filter all current students


##### Independent variables to consider:
For earning and unemployment
* hrlyearn: Usual hourly wages
* WKSUEM: All persons aged 16 to 69 and persons aged 70 and over who worked during the reference year

* USZGAP: Adjusted size of residence,
* AGEGP: Age group,
* MARST: Marital status,
* FWORKED: Flag - employed during refyear,
* ALFST: annual labour force status,
* ATINC: after tax income,
* CPPQPP: CPP/QPP benefits,
* CQPC: Pension plan contributions,3
* EARNG: earnings,
* MAJRI: Major source of income,
* OAS: Old age security pension,
* RSPWI: RRSP withdrawals,
* UIBEN: employment insurance benefits,
* HHSIZE: number of household members,
* DWLTYP: type of dwelling,
* DWTENR: ownership of dwelling,
* REPA: condition of dwelling

  + Any combination of these variables may be considered to quantify QOL 

## Analysis 

(*Main body of analysis, containing data summaries (tables/plots),  statistical analyses (estimations, tests), and modelling (regression, classification) *)

### Some Subsection Title

<!-- You can organize content in subsections -->

(*Plots should be created in R, and not imported from file*)

```{r, echo = FALSE, fig.width = 5, fig.heigh = 5}
tibble( t = seq(0, 2*pi, by=0.1) ) %>% 
  mutate( x = 16*sin(t)^3, 
          y = 13*cos(t)-5*cos(2*t)-2*cos(3*t)-cos(4*t) ) %>% 
  ggplot(aes( x = x, y = y)) + 
  geom_polygon( fill = "pink") +
  coord_fixed() 
```
```{r}
ngs = read.csv("ngs.csv")
cis = read.csv("cis.csv")
lfs = read.csv("lfs.csv")
ngs %>% head
cis %>% head
lfs %>% head
```
```{r}
library(ggplot2)
#the highest education effect on 1990 later
lfs_90 = lfs %>% mutate(educ90 = factor(educ90, levels=0:6, labels = c("N", "N", "N", "NF", "T", "T", "T"))) %>% mutate(pick = is.na(educ90)) %>% filter(pick == FALSE) %>% select(1:81) 
 lfs_90 %>% ggplot() + geom_boxplot(aes(x = educ90, y=hrlyearn)) + xlab("educ_status")
```
> Firstly we do a basic comparison of the hourly earning between the highest education level. The data gives 7 different types of education level, from entry level to the Master degree. We choose the education data of 1990 later to make sure its up to date and devide these people into three groups showing the level they engage in the post secondary education. 
>Based on the boxplot, it is clear that in average, most people who obatained a university degree are tend to have a higher hourly income. Also, the highest income that group of people can reach are much higher that two other groups. 
>Another thing that can be discovered from the graph is that though highly closed, the *NF* group(people who does attend but not finish the post secondary education), however, have the lowest income on average, and the also the highest income of that group is the lowest among that of the three groups. Which may shows that the paid back of taking university eucation is also related to whether taking the complete 4 years or more. 

>To make it easier to test for different age, we now combine NF and N(since their result are really close considering hrlyearn)

```{r}
lfs_90_N = lfs %>% mutate(educ90 = factor(educ90, levels=0:6, labels = c("N", "N", "N", "N", "T", "T", "T"))) %>% mutate(pick = is.na(educ90)) %>% filter(pick == FALSE) %>% select(1:81) %>% filter(educ90 == "N") 

lfs_90_T = lfs %>% mutate(educ90 = factor(educ90, levels=0:6, labels = c("N", "N", "N", "N", "T", "T", "T"))) %>% mutate(pick = is.na(educ90)) %>% filter(pick == FALSE) %>% select(1:81) %>% filter(educ90 == "T")

N_age = lfs_90_N %>% group_by(age_12) %>% summarise(avg_income_N = mean(hrlyearn, na.rm = TRUE))

T_age = lfs_90_T %>% group_by(age_12) %>% summarise(avg_income_T = mean(hrlyearn, na.rm = TRUE))

left_join(T_age, N_age, by = "age_12") %>% mutate(diff = avg_income_T - avg_income_N)

left_join(T_age, N_age, by = "age_12") %>% mutate(diff = avg_income_T - avg_income_N) %>% ggplot(aes(x = age_12, y = diff)) + geom_point( )
```

>Based on the data, we group the peoples age for 5 years each. It is clear that for all age group, owning a secondary eduaction degree is a guarantee of higher hourly earn.
>For more specific apporach, we then look at the difference of the difference of hourly earning between people who have degree and who not. The difference keep increasing till it reach the highest pointat group10, age between 60-64. This is mainly showing that the effect of posy secondary degree will become larger as age grows. While the data decrease after group10, is mainly due to the retirement and also back to years ago the post-secondary are not that necessary in 20th century.

```{r}
cis %>% select(hlev2g)
cis_new = cis %>% filter(hlev2g %in% c(1,2,3,4)) %>% mutate(hlev2g = factor(hlev2g, levels=1:4, labels = c("N", "N", "N", "T"))) %>% filter(wksuem %in% c(0:52)) 



cis_new %>% filter(wksuem != 0) %>% ggplot(aes(x = wksuem, fill = hlev2g))+geom_histogram()

```
>Also, we use the data from ngs to see if the post secondary degree would effect people on their unemployment length. Based on the graph above, though for almost each length there are both people with no degree and people with degree, but the people who don't have degree tend to have longer weeks of unemployment. 


#### Some Subsubsection Title

(*R output tables can be converted to .html with* `knitr::kable()`) 

```{r, echo = FALSE}
tibble( X = c("this", "is"), 
        Y = c("my", "table") ) %>% 
  knitr::kable()
```
<!-- 
for more help on RMarkdown see: https://rmarkdown.rstudio.com/lesson-15.html 
-->


## Summary 

(*Recap of your findings and conclusions*)



***  

## Appendix 

(*Everything here does not count towards page limit*)


