---
title: "STAA57 W19 - Project Proposal"
author:  "Team G?"
output: html_notebook
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

***

#### Research Questions

1. Specify which questions you will address.

The initial question of "Is university education worth it?" was rather broad so we wanted to find a clear concise way of simplifying it while still trying our best to answer it. One of the major factors about post secondary education in general is to try to find a better job than if you were to have skipped out on it entirely. So we wanted to delve deeper into the world of if getting a degree actually affects how much money you end up making in the future. The question we will be focusing our research on will be "Does a university degree actually translate to better pay in the future?" and then hopefully we can investigate "How fast will you be able to recoup your investment in post secondary education?".



2. Specify your data analysis plan

In our analysis we will be using a combination of education data and employment data. A good place to start will be to start looking at data that explains or shows how much people with all kinds of education backgrounds make. So we could filter the observations in the data set we use to see which individuals did receive a university degree and the ones that didn't. From there we could then look at the average of the salaries of the university graduates and compare it to the average of the salaries of workers who did not graduate and compare the values and try to draw a conclusion.We can also look at the employment of university graduates versus those who did not attend university to see if those individuals have higher employment. So from there, we can look calculate the unemployment rate of university graduates (doing the relevant calculations of course) and compare it to the unemployment of those who did not go to university and try to further support our argument.



#### Data Description

3. Specify your data sources and data collection methods.


4. Identify the variables/observations you will use for your analysis.

>
**National Graduate Survey (2013)**:
The following variables are key in analyzing our data. They will be used to filter, sort, and correlate data accordingly.
>
* CNDPERSP, total personal income from all sources in 2013
* JOBINCP, gross annual earnings from 2013
* DBTALGRD, debt of all loans at time of graduation (2009/10)
* DBTALINT, debt of all loans at time of interview (2013)
* DBTGSGRD, debt of government sponsored student loans at time of graduation (2009/2010)
* DBTGSINT, debt of government sponsored student loans at time of interview (2013)
* LFSTAT, labour force status in 2013 reference week

>
**Canadian Income Survey (2013)**:
Hundreds of variables are provided, but the variables key to our analysis are listed below.
>

* AGEGP, person’s age group as of December 31 of 2013
* HLEV2G, highest level of education of person
* STUDFTP, attended school/college/CEGEP/university during 2013
* EFATINC, after-tax income



#### Getting Started 

5. Load data into R.

```{r}
ngs = read.csv("ngs.csv")
cis = read.csv("cis.csv")
ngs %>% head
cis %>% head
```

6. Explore data 

```{r}
ngs %>% mutate(lfstat = (lfstat == 1)) %>%
  ggplot(aes(lfstat)) + geom_bar() + xlab("Currently Employed") + ylab("frequency")
```

```{r}
ngs %>% filter(lfstat == 1, jobincp != 99 & jobincp != 96) %>% mutate(jobincp = factor(jobincp, labels = c("0-9 999", "-19 999", "-29 999", "-39 999", "-49 999", "-59 999", "-69 999", "-79 999", "-89 999", "90 000+"))) %>% 
  ggplot(aes(jobincp)) + geom_bar() + xlab("Annual income") + ylab("frequency")
```

```{r}
ngs %>% filter(lfstat == 1, jobincp != 99 & jobincp != 96, dbtalgrd != 6 & dbtalgrd != 9) %>% 
  mutate(jobincp = factor(jobincp, labels = c("0-9 999", "-19 999", "-29 999", "-39 999", "-49 999", "-59 999", "-69 999", "-79 999", "-89 999", "90 000+"))) %>%
  group_by(jobincp) %>% summarise(owed = mean(dbtalgrd)) %>%
  ggplot(aes (jobincp, owed)) + geom_histogram(stat = "identity") + ylab("Amount owed (2 = 5k-10k)") + xlab("salary")
```

