---
title: "STAA57 W19 - Project Proposal"
author:  "Team G?"
output: html_notebook
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

***

#### Research Questions

1. Specify which questions you will address.


2. Specify your data analysis plan 


#### Data Description

3. Specify your data sources and data collection methods.


4. Identify the variables/observations you will use for your analysis.

>
**National Graduate Survey (2013)**:
The following variables are key in analyzing our data. They will be used to filter, sort, and correlate data accordingly.
>
* CNDPERSP, total personal income from all sources in 2013
* JOBINCP, gross annual earnings from 2013
* DBTALGRD, debt of all loans at time of graduation (2009/10)
* DBTALINT, debt of all loans at time of interview (2013)
* DBTGSGRD, debt of government sponsored student loans at time of graduation (2009/2010)
* DBTGSINT, debt of government sponsored student loans at time of interview (2013)
* LFSTAT, labour force status in 2013 reference week


#### Getting Started 

5. Load data into R.
>
```{r}
ngs = read.csv("ngs.csv")
#cis = read.csv("cis.csv")
ngs %>% head
#cis %>% head
```

6. Explore data 


