---
title: "STAA57 W19 - Project Proposal"
author:  "Team G?"
output: html_notebook
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

***

#### Research Questions

1. Specify which questions you will address.


2. Specify your data analysis plan 


#### Data Description

3. Specify your data sources and data collection methods.


4. Identify the variables/observations you will use for your analysis.

>
**National Graduate Survey (2013)**:
The following variables are key in analyzing our data. They will be used to filter, sort, and correlate data accordingly.
>
* CNDPERSP, total personal income from all sources in 2013
* JOBINCP, gross annual earnings from 2013
* DBTALGRD, debt of all loans at time of graduation (2009/10)
* DBTALINT, debt of all loans at time of interview (2013)
* DBTGSGRD, debt of government sponsored student loans at time of graduation (2009/2010)
* DBTGSINT, debt of government sponsored student loans at time of interview (2013)
* LFSTAT, labour force status in 2013 reference week


#### Getting Started 

5. Load data into R.

```{r}
ngs = read.csv("ngs.csv")
cis = read.csv("cis.csv")
ngs %>% head
cis %>% head
```

6. Explore data 

```{r}
ngs %>% mutate(lfstat = (lfstat == 1)) %>%
  ggplot(aes(lfstat)) + geom_bar() + xlab("Currently Employed") + ylab("frequency")
```

```{r}
ngs %>% filter(lfstat == 1, jobincp != 99 & jobincp != 96) %>% mutate(jobincp = factor(jobincp, labels = c("0-9 999", "-19 999", "-29 999", "-39 999", "-49 999", "-59 999", "-69 999", "-79 999", "-89 999", "90 000+"))) %>% 
  ggplot(aes(jobincp)) + geom_bar() + xlab("Annual income") + ylab("frequency")
```

```{r}
ngs %>% filter(lfstat == 1, jobincp != 99 & jobincp != 96, dbtalgrd != 6 & dbtalgrd != 9) %>% 
  mutate(jobincp = factor(jobincp, labels = c("0-9 999", "-19 999", "-29 999", "-39 999", "-49 999", "-59 999", "-69 999", "-79 999", "-89 999", "90 000+"))) %>%
  group_by(jobincp) %>% summarise(owed = mean(dbtalgrd)) %>%
  ggplot(aes (jobincp, owed)) + geom_histogram(stat = "identity") + ylab("Amount owed (2 = 5k-10k)") + xlab("salary")
```

